---
title: "Projects in Computational Modelling of Social Systems"
author: "David Garcia, Petar Jerčić, Jana Lasser <br><br> *TU Graz*"
date: "Computational Modelling of Social Systems"
output:
  xaringan::moon_reader:
    lib_dir: libs 
    css: [xaringan-themer.css, "libs/footer.css"]
    nature:
      beforeInit: ["libs/perc.js"]
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
---

```{r xaringan-themer, include=FALSE, warning=FALSE}
#This block contains the theme configuration for the CSS lab slides style
library(xaringanthemer)
library(showtext)
style_mono_accent(
  base_color = "#5c5c5c",
  text_font_size = "1.5rem",
  header_font_google = google_font("Arial"),
  text_font_google   = google_font("Arial", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
```

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

layout: true

<div class="my-footer"><span>David Garcia - Computational Modelling of Social Systems</span></div> 

---

# Project basics

- Participation in group presentation sessions (both as presenter and discussant): 25% of final grade

- Project report and codes: 25% of final grade

- Timeline:
  - 2.06: First project feedback exercise session
  - 9.06: Second project feedback exercise session and general Q&A session
  - 10.06: last chance to get any feedback from lecturers online via discord
  - 11.06: Deadline to register your group in Teach Center.
  - 23.06: First presentations session
  - 30.06: Second presentations session
  - 10.07: Project report deadline


---

# Projects objectives

With the projects, we want to evaluate the following skills and learning:

- Planning a computational project as in the exercises and course examples: model, simulation, analysis, interpretation

- Describing a model in detail, referring to other relevant models (especially if covered by the course)

- Ability to implement and code the model, showing it works as designed with visualizations and other outputs

- Assessing the project question with systematic analysis of simulations

---

# Project routes

A) Reproducing the analysis of a research paper
  - Example: Reproducing the results of Axelrod's culture model paper
  - Reflect on similarities or differences in your results
  - If paper is short, explore a small extra question or analysis
  
B) Extending a model from the course
  - Example: Neighbor overlap in the Barabasi-Albert model
    - Motivation: online we don't only interact with our friends in a direct manner, but also through common friends. Users with larger and more clustered neighborhoods will be less likely to leave
    - Analysis: measuring average neighbor overlap as network grows
    - Question: does overlap change with node age and network size?


---
# Project routes (II)

C) Designing and analyzing a new model
  - Example: privacy concerns in social resilience
    - Model probability to leave a social network increasing with too many friends
    - Study % of users leaving with value of parameter of that increase
    - Check previous work! Is this really a new model?
    - Can be the most challenging approach

** Data is not always necessary but desirable in some cases**
  - Data-driven simulations (e.g. using network data)
  - Comparing outcomes with empirical data (hard)
  - Quantitatively validating outcomes (harder)
  - Calibrating dynamics (hardest)
---

# Modeling project pitfalls

- **Not understanding your question**

  - Bad: *I want to model how people play Pokemon Go*
  - Good: *I want to understand the role of city size in Pokemon Go players using only two teams*
  - Risk: not knowing when you are done and if you are doing it right


- **Not searching for previous work**
  - Bad: *Here is the first model ever for fashion*
  - Good: *My model is similar to (cite,cite) in this and that aspect*
  - Risk: specifying the problem wrong or missing interesting parts
  
---

# Modeling project pitfalls (II)

- **Not being systematic**
  - Bad: *I made some simulations and they look like this*
  - Good: *Outcome measure X grows with parameter Y as shown in the boxplot*
  - Risks: Not having any substantial analysis and just anecdotes. Going back and forth between model and simulations in an endless loop.


- **Being too ambitious**
  - Bad: *I want to explain polarization*
  - Good: *I want to understand how the strength of social influence affects polarization of opinions in the social network of German politicians on Twitter*
  - Risks: Too much work and open paths, large models lead to complicated analyses. Might have too much previous work behind.


